---
title: "HW 5"
output:
  pdf_document: default
  html_document: default
---

# Problem 1
## (1)
Since we have a uniform prior the posterior is  
\begin{align*}
 p(\theta | y) &\propto I_{[10^{-5}, 10^5]}(\theta) \prod_{i=1}^{n} \exp\left(-\frac{(y_i-\theta)^2}{2}\right)\\
&\propto I_{[10^{-5}, 10^5]}(\theta)  \exp\left(-\frac{ n\theta^2+2\sum_{i=1}^{n}y_i\theta}{2}\right)\\
&\propto I_{[10^{-5}, 10^5]}(\theta)  \exp\left(-\frac{(\theta-\bar{y})^2}{2n^{-1}}\right).
\end{align*}
It's $N(\bar{y}, n^{-1})$ truncated in $[10^{-5}, 10^5]$. So the posterior predictive distribution is 
\begin{align*}
p(y^{rep} | y) &\propto \int p(y^{rep}|\theta)p(\theta | y) \;d \theta\\
&\propto \int \exp\left(-\frac{(y^{rep}-\theta)^2}{2}\right)\exp\left(-\frac{(\theta-\bar{y})^2}{2n^{-1}}\right)I_{[10^{-5}, 10^5]}(\theta) \;d \theta
\end{align*}

```{r}
set.seed(0)

posterior <- rnorm(1e4, mean=5.1, sd=100^{-1/2})
posterior <-posterior[(posterior >= 1e-5 & posterior <= 1e5)]

T_rep <- c()
for (theta in posterior) {
  rep <- rnorm(100, mean=theta, sd=1)
  T_rep <- c(T_rep, max(rep))
}

hist(T_rep)

cat('p-value is', mean(T_rep >= 8.1))
```

## (2)
The prior predictive distribution is 
\begin{align*}
p(y^{rep}) &\propto \int p(y^{rep}|\theta)p(\theta ) \;d \theta\\
&\propto \int \exp\left(-\frac{(y^{rep}-\theta)^2}{2}\right)I_{[10^{-5}, 10^5]}(\theta) \;d \theta\\
&\propto P_{N(y^{rep}, 1)}\left(X \in [10^{-5}, 10^5]\right).
\end{align*}
```{r}
prior <- runif(1e4, min=1e-5, max=1e5)

T_rep_prior <- c()
for (theta in prior) {
  rep <- rnorm(100, mean=theta, sd=1)
  T_rep_prior <- c(T_rep_prior , max(rep))
}

hist(T_rep_prior )

cat('p-value is', mean(T_rep_prior >= 8.1))
```


## (3)
Obviously the prior does not contain any information of data $y$, and it's not very informative with pre-existing knowledge. So just using the prior we won't be able to replicate data $y$.  

## Problem 2

According to Chapter $5$ the tumor rate $\theta$ has a Beta prior with parameters $\alpha=1.4$ and $\beta=8.6$ estimated using historical data. The posterior is $\text{Beta}(5.4, 18.6)$. 

Define test quantities as follows:

* The max tumor rate. $T_1(y) = \max(y/n)$. 
* Measure the fluctuation. $T_2(y) = (n-1)^{-1}\sum(y_i/n_i-\overline{y_i/n_i})^2$. 
* How many groups have $0$ tumor rat. $T_3(y) = \sum I(y_i = 0)$. 

```{r}
n <- c(20,20,20,19,18,18,18,18,25,20,10,49,20,13,48,20,48,19,23,19,22,20,20,20,17,20,20,24,23,20,19,46,27,50,20,20,19,19,22,20,20,20,20,19,20,20,20,20,17,49,20,20,46,49,52,47,19,19,19,19,20,20,47,20,20,20,20,20, 46,24,14)
y <- c(0, 0, 0, 0, 0, 0, 1, 1, 2, 2, 1, 5, 3, 2, 9, 4, 10, 4, 6, 5, 6, 0, 0, 0, 0, 1, 1, 2, 2, 2, 2, 5, 3, 10, 4, 4, 4, 4, 5, 6, 6, 6, 0, 0, 1, 1, 2, 2, 2, 7, 4, 4, 11, 12, 16, 15, 0, 0, 1, 1, 2, 2, 7, 3, 4, 4, 5, 5,15, 9, 4)

T1 <- max(y/n)
T2 <- var(y/n)
T3 <- sum(y/n == 0)
  

T1_rep <- c()
T2_rep <- c()
T3_rep <- c()
for (i in 1:1e4) {
  theta <- rbeta(71, 5.4, 18.6)
  rep <- c()
  for (j in 1:71) {
    rep <- c(rep, rbinom(1, size=n[j], theta[j])/n[j])
  }
  T1_rep <- c(T1_rep, max(rep))
  T2_rep <- c(T2_rep, var(rep))
  T3_rep <- c(T3_rep, sum(rep == 0))
}

hist(T1_rep)
cat('p-value for T1 is', mean(T1_rep >= T1))
```

```{r}
hist(T2_rep)
cat('p-value for T2 is', mean(T2_rep >= T2))
```

```{r}
hist(T3_rep)
cat('p-value for T3 is', mean(T3_rep >= T3))
```





